---
title: "Predicting reproductive toxicity: a model for PFAS-like chemical structures"
subtitle: "BMIN503/EPID600 Final Project"
author: "Andrea Andress Huacachino"
format:
  html:
    toc: true
    toc-location: left
editor: visual
embed-resources: true
execute: 
  warning: false
  message: false
---

------------------------------------------------------------------------

### Overview

Per- and polyfluoroalkyl substances (PFAS) are a class of chemicals that pollute our environment. They are ubiquitous, and nearly in 99% of all humans on earth. Unfortunately, some members have been associated to various health outcomes, including reproductive health. Using supervised machine learning, the objective of the project is to predict the reproductive toxicity of PFAS members by characterizing their chemotypes and relating them to known biological outcomes. The data for the project was obtained from Tox 21 and  CompToxAI. Dr. Joseph Romano assisted with obtaining the data while Dr. Trevor Penning assisted by redefining the assays chosen as the biological endpoint. 

Due to growing scrutiny by civilian and governmental organizations, many legacy PFAS have been pressured to be rolled back. However, in order to maintain profits, producers have been creating new PFAS. Many of the new "next generation" PFAS take advantage of the lack of toxicological data available to claim to be safer, despite no studies suggesting safety. The necessity for predicting toxicity from PFAS structure knowledge is then evident. The significance of my project is to create a machine learning model to predict toxicity in novel PFAS for the ultimate goal to prioritize further studies and to help guide PFAS production towards a less toxic future. 

All data is publicly available on my github account, TOX21, and CompToxAI.

GitHub repository:

### Introduction

Describe the problem addressed, its significance, and some background to motivate the problem.

Explain why your problem is interdisciplinary, what fields can contribute to its understanding, and incorporate background related to what you learned from meeting with faculty/staff.

### Methods

Describe the data used and general methodological approach. Subsequently, incorporate full R code necessary to retrieve and clean data, and perform analysis. Be sure to include a description of code so that others (including your future self) can understand what you are doing and why.

**Loading in Packages**

```{r}
#| eval: FALSE

#install.packages("kernlab")
#install.packages("randomForest")
#install.packages("glmnet")
#install.packages("tidymodels")
#install.packages("dotwhisker")
#install.packages("vip")




library(tidyverse)
library(kernlab)
library(randomForest)
library(glmnet)
library(tidymodels)
tidymodels_prefer()
library(dotwhisker)
library(vip)

```

**Reading in Files**

All files have been retrieved from [CompToxAI](https://comptox.ai/data.html) and [Tox21](https://tripod.nih.gov//tox21/pubdata/). Once downloaded onto my computer, I uploaded the files on my github in order for the data to be publicly available in an ordered csv file. Each file contains structural data pertaining to the chemical entity and toxicity by showing if it interacted with the target or not. First column is the chemical ID, then columns 2-167 are structural data of the chemical, finally column 168 is if it interacted with the target, 0 for nontoxic, 1 for toxic.

```{r}

#Descriptors in file name: protocol, target, cell type, agonist or antagonist
 

#Androgen receptor (AR) based assays 

ar_bla_agonist_p1 <- read.csv('https://raw.githubusercontent.com/AndreaHuacachino/BMIN503_Final_Project/master/andrea_tox21-ar-bla-agonist-p1.csv', header = TRUE)

ar_bla_antagonist_p1 <- read.csv('https://raw.githubusercontent.com/AndreaHuacachino/BMIN503_Final_Project/master/andrea_tox21-ar-bla-antagonist-p1.csv', header = TRUE)

ar_mda_kb2_luc_agonist_p1 <- read.csv('https://raw.githubusercontent.com/AndreaHuacachino/BMIN503_Final_Project/master/andrea_tox21-ar-mda-kb2-luc-agonist-p1.csv', header = TRUE)

ar_mda_kb2_luc_agonist_p3 <- read.csv('https://raw.githubusercontent.com/AndreaHuacachino/BMIN503_Final_Project/master/andrea_tox21-ar-mda-kb2-luc-agonist-p3.csv', header = TRUE)

ar_mda_kb2_luc_antagonist_p1 <- read.csv('https://raw.githubusercontent.com/AndreaHuacachino/BMIN503_Final_Project/master/andrea_tox21-ar-mda-kb2-luc-antagonist-p1.csv', header = TRUE)

ar_mda_kb2_luc_antagonist_p2 <- read.csv('https://raw.githubusercontent.com/AndreaHuacachino/BMIN503_Final_Project/master/andrea_tox21-ar-mda-kb2-luc-antagonist-p2.csv', header = TRUE)



#Estrogen Receptor Assays

er_bla_agonist_p2 <- read.csv('https://raw.githubusercontent.com/AndreaHuacachino/BMIN503_Final_Project/master/andrea_tox21-er-bla-agonist-p2.csv', header = TRUE)

er_bla_antagonist_p1 <- read.csv('https://raw.githubusercontent.com/AndreaHuacachino/BMIN503_Final_Project/master/andrea_tox21-er-bla-antagonist-p1.csv', header = TRUE)

er_luc_bg1_4e2_agonist_p2 <- read.csv('https://raw.githubusercontent.com/AndreaHuacachino/BMIN503_Final_Project/master/andrea_tox21-er-luc-bg1-4e2-agonist-p2.csv', header = TRUE)

er_luc_bg1_4e2_agonist_p4 <- read.csv('https://raw.githubusercontent.com/AndreaHuacachino/BMIN503_Final_Project/master/andrea_tox21-er-luc-bg1-4e2-agonist-p4.csv', header = TRUE)

er_luc_bg1_4e2_antagonist_p1 <- read.csv('https://raw.githubusercontent.com/AndreaHuacachino/BMIN503_Final_Project/master/andrea_tox21-er-luc-bg1-4e2-antagonist-p1.csv', header = TRUE)

er_luc_bg1_4e2_antagonist_p2 <- read.csv('https://raw.githubusercontent.com/AndreaHuacachino/BMIN503_Final_Project/master/andrea_tox21-er-luc-bg1-4e2-antagonist-p2.csv', header = TRUE)

# Estrogen Receptor subtype: beta
erb_bla_antagonist_p1 <- read.csv('https://raw.githubusercontent.com/AndreaHuacachino/BMIN503_Final_Project/master/andrea_tox21-erb-bla-antagonist-p1.csv', header = TRUE)

erb_bla_p1 <- read.csv('https://raw.githubusercontent.com/AndreaHuacachino/BMIN503_Final_Project/master/andrea_tox21-erb-bla-p1.csv', header = TRUE)



#PPAR-gamma Assays

pparg_bla_agonist_p1 <- read.csv('https://raw.githubusercontent.com/AndreaHuacachino/BMIN503_Final_Project/master/andrea_tox21-pparg-bla-agonist-p1.csv', header = TRUE)

pparg_bla_antagonist_p1 <- read.csv('https://raw.githubusercontent.com/AndreaHuacachino/BMIN503_Final_Project/master/andrea_tox21-pparg-bla-antagonist-p1.csv', header = TRUE)


head(ar_bla_agonist_p1)
```

Making random forest engine

```{r}


rf_cls_spec <- 
  rand_forest(trees = 100, min_n = 5) |> 
  set_engine("randomForest", importance = TRUE) |>
  set_mode("classification")
rf_cls_spec


#ar_bla_agonist_p1

#making dataset include factors
ar_bla_agonist_p1.2 <- ar_bla_agonist_p1 |>
  mutate(target = factor(target, levels = c(0, 1), 
                           labels = c("nontoxic","toxic")))


ar_bla_agonist_p1.2$target


#Making model
ar_bla_agonist_p1.rf_cls_fit <- rf_cls_spec |>
  fit(target ~ ., data = ar_bla_agonist_p1.2)


ar_bla_agonist_p1.rf_cls_fit

ar_bla_agonist_p1.rf.pred.values <- bind_cols(
  truth = ar_bla_agonist_p1.2$target,
  predict(ar_bla_agonist_p1.rf_cls_fit, ar_bla_agonist_p1.2),
  predict(ar_bla_agonist_p1.rf_cls_fit, ar_bla_agonist_p1.2, type = "prob"))

#I keep getting Error in predict.randomForest(object = object$fit, newdata = new_data) : New factor levels not present in the training data SAD 




autoplot(roc_curve(ar_bla_agonist_p1.rf.pred.values, 
                   truth, .pred_died))



#cross validation stuff

rf_workflow <-
  workflow() |>
  add_model(rf_cls_spec) |>
  add_formula(survived ~ .)

set.seed(1234)
rf_fit_cv <-
  rf_workflow |>
  fit_resamples(titanic.training.folds, 
                control = control_resamples(save_pred = TRUE))

collect_metrics(rf_fit_cv)

rf_fit_cv |>
  collect_predictions() |>
  roc_curve(survived, .pred_died) |>
  autoplot()
```

### Results

Describe your results and include relevant tables, plots, and code/comments used to obtain them. End with a brief conclusion of your findings related to the question you set out to address. You can include references if you'd like, but this is not required.

**Conclusion**
